```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1.Read and check data
```{r}
#read and check data

marketing<- read.csv("marketing.csv")   

dim(marketing)
head(marketing)

```

# 2.Check for NA values
```{r}
                                    
any(is.na(marketing))                #If there´s any NA

colSums(is.na(marketing))            #NA for each var

#----------------------------------------------------------------------------------
# ---- Missingness overview ----

n<-nrow(marketing)  #number of rows in dataset               

na_count<-colSums(is.na(marketing))   #NA number for each vari
na_pct<-round(na_count / n*100, 2)    #% as in total, keep 2 decimals

na_sum<-data.frame(                  #build a summary table
  variable=names(na_count),
  na_count=as.integer(na_count),
  na_pct=na_pct)

na_sum[na_sum$na_count>0, ]   #only show NA vari
                               
```
## Initially, We examined the missing values.It shows that all 79 missing observations are located in the variable Income, representing approximately 1% of the dataset. No other variables contain missing values.


```{r}
#----------------------------------------------------------------------------------
#Income missingness diagnostics

marketing$Income_missing<-is.na(marketing$Income)          #logical indicator NA-T other-F

vars_cat<-c("Gender","CampaignChannel","CampaignType","Conversion")

for (var in vars_cat) {                                          #Loop variables f missing rate Income
  cat("\n Income missing by", var,"\n")                          #Print header label
  tab<-table(marketing[[var]],marketing$Income_missing)        #contingency table:rows:categories,columns:Income_NA F/T
  print(round(prop.table(tab,1)[,"TRUE"],4))}                  #missing rate within each group

```
## The 1% NA in Income is fairly balanced across categories.
Only Age shows a small but statistically significant difference.
Overall, Income NA is minor and unlikely to affect the analysis.


```{r}
#Compare numeric variables by Income missingness

vars_num<-c("Age", "AdSpend")                                                   #Select num-variables t compare

for (var in vars_num) {                                                        #Loop each num-variable
  cat("\n", var, "\n")                                                         #header
  print(tapply(marketing[[var]],
               marketing$Income_missing,                                      #separately for Inc-miss T/F
               summary))}                                                     #Compute summary statistics

chisq.test(table(marketing$CampaignChannel,
                 marketing$Income_missing))                                #Chi-square-test for categorical association

```
## Income missingness slightly related to Age, but shows no strong association with AdSpend or CampaignChannel.


# 3.Check for duplicate values
```{r}
#chek dup amount
dupid_count<-sum(duplicated(marketing$CustomerID))        #Check duplicate CustomerID
dupid_count

duprow_count<-sum(duplicated(marketing))                  #Check fully-dup rows
duprow_count

#check if the same dup
dupid_rows<-which(duplicated(marketing$CustomerID))       #dupID rows
dupfull_rows<-which(duplicated(marketing))                #dup full rows
all(dupid_rows==dupfull_rows)                          #compare if same

marketing<-marketing[!duplicated(marketing), ]             #remove f-dup rows

dim(marketing)                                              #new dimension

```
## We found 200 duplicated CustomerID values and by further inspection show these 200 exactly are the 200 fully duplicated rows.Therefore, we remove them, and after removing the duplicated rows the dataset contains 8000 observations and 19 variables.


# 4.Validity & consistency checks
```{r}
#Income IQR outlier
Q1<-quantile(marketing$Income,0.25,na.rm=TRUE)   #25th percentile
Q3<-quantile(marketing$Income,0.75,na.rm=TRUE)   #75th percentile
IQR_val<-IQR(marketing$Income,na.rm=TRUE)         #interquartile range

low_fence<-Q1-1.5 * IQR_val
up_fence<-Q3+1.5 * IQR_val                     #up&low boundary

sum(marketing$Income<low_fence|marketing$Income>up_fence,na.rm=TRUE)     #Count num of outliers

marketing$Income[!is.na(marketing$Income)&
                   (marketing$Income<low_fence|marketing$Income>up_fence)]   #show outlier values(exclud NA)

marketing$Income[marketing$Income==1e+21]<-NA                            #Replace all 1e+21 with NA
sum(marketing$Income==1e+21,na.rm=TRUE)                              #Verify no 1e+21 remain

```
## The IQR method was used to identify potential outliers in Income. Two extreme values both equal to 1e+21, which is not a plausible income value and was therefore recoded as NA.


```{r}
#Summary for key numeric variables
summary(marketing[,c("Age","Income","AdSpend","ClickThroughRate","ConversionRate","WebsiteVisits",
                     "PagesPerVisit","TimeOnSite","SocialShares","EmailOpens","EmailClicks",
                     "PreviousPurchases","LoyaltyPoints")])

#ConversionRate must between 0-1
marketing$ConversionRate[marketing$ConversionRate<0|marketing$ConversionRate>1]<-NA   

summary(marketing$ConversionRate)

```
## The summary revealed the ConversionRate contained values outside the valid range 0-1. So the values were treated as errors and recoded as NA.

```{r}
#Export cleaned dataset
write.csv(marketing,"marketing_cleaned Pt1.csv",row.names=FALSE)


## Handover to Part 2
## Data cleaning and validity checks have been completed.
## All invalid entries have been recoded as NA.
## The dataset is now ready for exploratory analysis and modelling.
## Please use marketing_cleaned.csv for Part 2 analysis.

```


